The Splashback Radius of Planck-SZ clusters

### 2018-11-15 Paper_new/RSP_paper_new.tex
minor edits

\begin{abstract}
We present evidence for the existence of the splashback radius in
galaxy clusters selected using the Sunyaev-Zeldovich effect.  We show
that the deprojected cross-correlation of galaxy clusters found in the
Planck survey with galaxies detected photometrically in the Pan-STARRS
survey, shows a sharp steepening feature (a logarithmic slope steeper
than $-3$), which we associate with the splashback radius. We infer
the three-dimensional splashback radius for the SZ cluster sample to
be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$ $\mpch$, where the cluster sample
has an average halo mass $M_{\rm 500c}=3.0\times10^{14}\msunh$ at an
average redshift of $z=0.18$. The inferred value of the splashback
radius appears consistent with the expected location for dark matter halos
in the standard cold dark matter paradigm. However, given the limited
precision of our measurements, we cannot conclusively rule out the
smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius
does not depend upon the galaxy magnitude for galaxies fainter than
$M_i-5\log h=-19.44$, and is present at a consistent location in
galaxy populations divided by color. The presence of the splashback
radius in the star-forming galaxy population could potentially be used
to put lower limits on the quenching timescales for galaxies. However,
these results need further verification with deeper datasets, as
photometric errors could cause contamination of the star-forming
galaxy sample by quenched galaxies.

@@

We restrict ourselves to the redshift range $0.03 \leq z \leq 0.33$ in order to
have a similar redshift range used in \citet{more2016detection}. The mass and
the redshift distribution of the Planck clusters we use is shown in
Fig.~\ref{fig:planck_summary}. Due to the larger beam size of the Planck
satellite, the cluster positions as reported in the catalog may be
mis-centered from the true centers of the galaxy clusters. We perform
a visual inspection of Pan-STARRS images taken around the detected
galaxy clusters in order to locate the nearest, brightest cluster
galaxy (BCG). We regard the position of the BCG as the true cluster
location, under the assumption that the BCG is located at the true,
gravitational center. Due to this crude assumption, we additionally 
study the effects of a possible, remaining mis-centering of the 
cluster positions in our model for the two-dimensional correlation 
function (see Appendix~\ref{sec:mis-centering}).

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \cdot 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

Given the flux limited galaxy catalog that we use, we expect to
observe more correlated galaxies in galaxy clusters that lie closer
to us, but with much fainter absolute magnitudes. To avoid
such biases with redshift of the clusters we restrict ourselves to
galaxies with absolute magnitudes brighter than a certain magnitude
limit, which depends on the depth of the used galaxy catalog. We
make the assumption that the galaxies reside at the redshift of the
cluster in question. The used magnitude limits are (-19.44, -18.94
and -18.44) for the catalogs PS 21, PS 21.5 and PS 22, respectively.

@@

In our analysis, the same galaxy may be considered red or blue depending upon
the cluster redshift under consideration. Our method avoids the use of
uncertain photometric redshifts to derive k-corrections
\citep[cf.][]{baxter2017halo}. We will study the cross-correlations to derive
the splashback radii for these galaxy populations separately, as well as the
fractional contributions of the red and the blue galaxies to the total density
profile as a function of radius. The caveats in the interpretation of these
results are mentioned in the Appendix~\ref{sec:Errors}.

@@

magnitude brighter than $M_{\rm i}-5\log h=-19.44$. The 68\%

@@

We caution however that the cut we use to define our samples was
defined based on spectroscopic galaxies, which tend to be brighter
than the typical galaxies we use for the cross-correlation. As we use
galaxies at fainter magnitudes for the cross-correlation, the
photometric errors increase dramatically and can scatter many more of
the red galaxy population into the blue galaxy sample. We report on
our investigations in Appendix~\ref{sec:Errors} as a cautionary tale,
and to guide future efforts to establish the splashback radius in the
blue galaxy population. We show that given the current state of the
data we cannot exclude a contamination of the correlation function of
blue galaxies from the red galaxy population and appropriate caution
be warranted in the interpretation or use of the results derived using
the color separated galaxy populations.

@@

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
the Appendix. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam (HSC) and the Dark Energy Survey (DES) would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope (LSST) would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
    \includegraphics[width= 0.5\textwidth]{color_vs_z.pdf}
\caption{Scatter plot showing the $(g_{\mathrm{PS}}-r_{\mathrm{PS}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}

@@

\begin{figure*}
\centering{
    \includegraphics[width= \textwidth]{mag_errors.pdf}
\caption{Shown are the photometric errors of the Pan-STARRS galaxies versus
their magnitude in each of the used band. The dashed curves indicate the running
median of the photometric error.}
   \label{fig:mag_errors} 
}
\end{figure*}

>>>>

\begin{abstract}
We present evidence for the existence of the splashback radius in
galaxy clusters selected using the Sunyaev-Zeldovich effect.  We show
that the deprojected cross-correlation of galaxy clusters found in the
Planck survey with galaxies detected photometrically in the Pan-STARRS
survey, shows a sharp steepening feature (a logarithmic slope steeper
than $-3$), which we associate with the splashback radius. We infer
the three-dimensional splashback radius for the SZ cluster sample to
be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$ $\mpch$, where the cluster sample
has an average halo mass $M_{\rm 500c}=3.0\times10^{14}\msunh$ at an
average redshift of $z=0.18$. The inferred value of the splashback
radius appears consistent with the expected location for dark matter halos
in the standard cold dark matter paradigm. However, given the limited
precision of our measurements, we cannot conclusively rule out the
smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius
does not depend upon the galaxy magnitude for galaxies fainter than
$M_i-5\log h=-19.44$, and is present at a consistent location in
galaxy populations divided by color. The presence of the splashback
radius in the star-forming galaxy population could potentially be used to put
lower limits on the quenching timescales for galaxies. We can marginally rule
out the contamination of the star-forming galaxy sample by quenched galaxies,
but the results would need further verification with deeper datasets.

@@

We restrict ourselves to the redshift range $0.03 \leq z \leq 0.33$ in order to
have a similar redshift range used in \citet{more2016detection}. The mass and
the redshift distribution of the Planck clusters we use is shown in
Fig.~\ref{fig:planck_summary}. Due to the larger beam size of the Planck
satellite, the cluster positions as reported in the catalog may be
mis-centered from the true centers of the galaxy clusters. We perform
a visual inspection of Pan-STARRS images taken around the detected
galaxy clusters in order to locate the nearest, brightest cluster
galaxy (BCG). We regard the position of the BCG as the true cluster
location, under the assumption that the BCG is located at the true,
gravitational center. Due to this assumption, we additionally 
study the effects of a possible, remaining mis-centering of the 
cluster positions in our model for the two-dimensional correlation 
function (see Appendix~\ref{sec:mis-centering}).

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \times 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

Given the flux limited galaxy catalog that we use, we expect to
observe more correlated galaxies in galaxy clusters that lie closer
to us, but with much fainter absolute magnitudes. To avoid
such biases with redshift of the clusters we restrict ourselves to
galaxies with absolute magnitudes brighter than a certain magnitude
limit, which depends on the depth of the used galaxy catalog. We
make the assumption that the galaxies reside at the redshift of the
cluster in question. The magnitude limits we use are -19.44, -18.94 and -18.44
and name the corresponding catalogs PS 21, PS 21.5 and PS 22, respectively.

@@

In our analysis, the same galaxy may be considered red or blue depending upon
the cluster redshift under consideration. Our method avoids the use of
uncertain photometric redshifts to derive k-corrections
\citep[cf.][]{baxter2017halo}. We will study the cross-correlations to derive
the splashback radii for these galaxy populations separately. The caveats in
the interpretation of these results due to photometric errors are discussed in the
Appendix~\ref{sec:Errors}.

@@

magnitude fainter than $M_{\rm i}-5\log h=-19.44$. The 68\%

@@

We caution however that the cut we use to define our samples was
defined based on spectroscopic galaxies, which tend to be brighter
than the typical galaxies we use for the cross-correlation. As we use
galaxies at fainter magnitudes for the cross-correlation, the
photometric errors increase dramatically and can scatter many more of
the red galaxy population into the blue galaxy sample. We report on
our investigations in Appendix~\ref{sec:Errors} as a cautionary tale,
and to guide future efforts to establish the splashback radius in the
blue galaxy population. We show that given the current state of the
data we can marginally exclude a contamination of the correlation
function of blue galaxies from the red galaxy population but
appropriate caution is warranted in the interpretation or use of the
results derived using the color separated galaxy populations.

@@

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
the Appendix~\ref{sec:Errors}. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam (HSC) and the Dark Energy Survey (DES) would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope (LSST) would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
\centering{
    \includegraphics[width= 0.5\textwidth]{color_vs_z.pdf}}
\caption{Scatter plot showing the $(g_{\mathrm{PS}}-r_{\mathrm{PS}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}

@@

\begin{figure*}
\centering{
    \includegraphics[width= \textwidth]{mag_errors.png}
\caption{Shown are the photometric errors of the Pan-STARRS galaxies versus
their magnitude in each of the used band. The dashed curves indicate the running
median of the photometric error.}
   \label{fig:mag_errors} 
}
\end{figure*}
